---
title: 'Capston Project: Exploratory Data Analysis Milestone'
author: "Bojan Brajkovic"
date: "21/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction to Analysis

The goal of this project is just to display that student is used to working with the data and is on track to create prediction algorithm. Report, should be submitted in RPubs, that explains exploratory analysis and goals for the eventual app and algorithm. This document should be concise and explain only the major features of the data identified and briefly summarize plans for creating the prediction algorithm and Shiny app in a way that would be understandable to a non-data scientist manager. Tables and plots should be used to illustrate important summaries of the data set. 

The motivation for this project is to: 
1. Demonstrate that you’ve downloaded the data and have successfully loaded it in.
2. Create a basic report of summary statistics about the data sets.
3. Report any interesting findings that you amassed so far.
4. Get feedback on your plans for creating a prediction algorithm and Shiny app.

##### Review criterialess
1. Does the link lead to an HTML page describing the exploratory analysis of the training data set?
2. Has the data scientist done basic summaries of the three files? Word counts, line counts and basic data tables?
3. Has the data scientist made basic plots, such as histograms to illustrate features of the data?
4. Was the report written in a brief, concise style, in a way that a non-data scientist manager could appreciate?

### Loading and downloading the data

##### Data download
``` {r}
if (!file.exists("./data/final")) {
  if (!file.exists("./data/Coursera-SwiftKey.zip")) {
    download.file("https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip",
                  destfile = "Coursera-SwiftKey.zip", quiet=TRUE)
  }
  
  unzip("Coursera-SwiftKey.zip", exdir = "./data");
}
```

##### Data Load
```{r}
setwd("~/Desktop/Capstone")
twitter <- readLines(connection <- file("data/final/en_US/en_US.twitter.txt"), encoding = "UTF-8", skipNul = TRUE)
close(connection)
blogs <- readLines(connection <- file("data/final/en_US/en_US.blogs.txt"), encoding = "UTF-8", skipNul = TRUE)
close(connection)
news <- readLines(connection <- file("data/final/en_US/en_US.news.txt"), encoding = "UTF-8", skipNul = TRUE)
close(connection)
rm(connection)
```

